---
title: "Social and economic networks"
subtitle: "Exam. Part 2"
author: "Juan José Martín"
output:
  html_document:
    number_sections: false
    highlight: tango
    toc: yes
    df_print: paged
editor_options: 
  chunk_output_type: console
---



<style type="text/css">

  body {
    background-color: #f6f7fd; 
  }
  
  a:link {
    color: #0174DF;
  }
  
  code.r {
    font-size: 14px;
  } 
  
  div pre {
    background-color:#E0ECF8;
  }
  pre {
    font-size: 14px 
  }
 
  p {
    text-align: justify;
  }
 
    
  h1, h2, h3, h4, h5, h6 {
    color: #737aaa;
  }

  th {  
    background-color:#737aaa;
    color: #FAFAFA;
    padding:5px;
  }
  
  td {
    font-size: 11.5pt;
  } 
  
  tr:nth-child(even){
    background: white;
  }
  
  tr:nth-child(odd){ 
    background-color: #EFF8FB;
  }
</style>

***  


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=8, fig.align="center", echo=TRUE, warning=FALSE, message=FALSE)
library(centiserve)
library(igraph)
```


**1) In a single sentence, what is the main question addressed in this paper? Why do the authors consider this question relevant and worth answering? After reading the paper, do you consider their answer relevant? Why?**

En el artículo se propone una nueva definición del concepto de triángulos y de un nuevo coeficiente de clustering generalizado ponderado.

Los autores lo consideran relevante porque con este nuevo coeficiente de clustering generalizado ponderado, se puede conocer la estructura de una red ponderada cuando faltan uniones entre nodos (triángulos indirectos).



<br>

**2) How does this paper relate to previous work about the question addressed in it?**
  
Se empieza mencionando varios artículos sobre medidas y centralidades para redes complejas. Posteriormente, se nombran artículos donde se explica el concepto de *local cluster coefficient* y un nuevo modelo combinatorio para detectar clusters que tengan en cuenta los datos individuales de una unidad de población y los datos que describen las conexiones entre dichas unidades. Finalmente, se mencionan diferentes análisis del coeficiente de clustering sobre redes ponderadas.

<br>

**3) What are the main definitions in this paper?**

La principal definición que proporciona el artículo en la del coeficiente de clustering generalizado ponderado. No obstante, para ello, definen dos nuevos tipos de triángulos alrededor de un nodo *i*:

* El primero está compuesto por dos nodos no unidos que son adyacentes a *i*.

* El segundo está compuesto por una trayectoria de distancia dos desde *i* tal que *i* no está unido con el nodo extremo de la trayectoria.

<br>

**4) On your behalf, what are the strongest contribution and the greatest drawback of this paper (if any)?**



<br>

**5) Do you consider the numerical experiments provided in Section 5 suitable? (I’m not referring to the networks analyzed, but to the computations performed on them.) What conclusions do you extract from these experiments? Do they agree with the authors’ conclusions? What other experiments (on the same networks) would you perform?**



<br>

**6) In their experiments, the authors mostly consider cases when $\alpha = 0$ or $\beta = 0$. How would you choose $\alpha$ and $\beta$ so that only really meaningful pseudo-triangles were considered?**

Los autores definen *T2* y *T3* como pseudo-triángulos porque, aunque no estén contenidos en *G*, pueden ser vistos como triángulos bajo las condiciones de los pesos de las dos aristas. Además, asumen $\alpha = \beta = 0$ para que el número de *T2* y *T3* alrededor de cada nodo no dependa de la elección de *F*.

<br>

**7) The authors define their generalized clustering coefficient only for undirected networks. How would you generalize it in a sound way to directed networks? Provide a definition that makes sense on some class of weighted directed networks of your choice with some specific meaning of its links (for instance, non-symmetric friendship networks, trade networks, citation networks with weights the number of times an author has cited another author, ...)**


<br>

**8) At the end of the day, the final value of the generalized clustering coefficient does not depend on the specific weights of the edges in the triangles and "generalized triangles": each such a triangle only counts 0 or 1 in the definition. Can you think of some sound way to use the weights to "weight" the triangles and "generalized triangles"" in the definition? How could your definition change the experimental results explained in the paper?**


<br>

**9) How is the generalized clustering coefficient related to the Strong Triadic Closure Property? Can you find a way to test whether a network satisfies the STCP using this coefficient (together with other indices, of course)?**


<br>

**Now, some technical questions:**

<br>

**10) In page 4 the authors say “Our generalized clustering coefficient has a further very relevant property: it assumes unitary value in several situations and not only when the graph is a clique.” Is it true that if an unweighted graph has clustering coefficient 1, then it is a clique? And if we assume moreover that the graph is connected? Justify your answer.**

Un clique es un grafo no dirigido *G* tal que para todo par de vértices *V*, existe una arista que las conecta. No obstante, para que un grafo tenga su coeficiente de clustering igual a 1, todo subgrafo del grafo *G* ha de tener coeficiente de clustering igual a 1, por lo que podría existir un grafo con varios subgrafos que dieran un coeficiente de clustering igual a 1, pero este no sería un clique. Por ejemplo, para el siguiente grafo

```{r}
g = graph_from_literal(1---2, 1---3, 1---4, 1---5, 2---3, 2---4, 2---5, 3---4, 3---5, 4---5, 6---7, 7---8, 8---6)
plot(g)
```

<br>

su coeficiente de clustering es igual a 1, pero no es un clique. Por lo tanto, no es cierto que cuando un grafo no ponderado *G* tiene el coeficiente de clustering igual a 1, entonces *G* es un clique.

```{r}
transitivity(g, type="average") # Clustering coefficient
```

<br>

No obstante, ante un grafo *G* conexo y con coeficiente de transitividad igual a 1, sí que sería un clique, ya que  todo par de vértices estaría unido con una arista, sino no cumpliría con el coeficiente de clustering igual a 1.

```{r}
g = graph_from_literal(1---2, 1---3, 1---4, 1---5, 2---3, 2---4, 2---5, 3---4, 3---5, 4---5)
plot(g)
```

```{r}
transitivity(g, type="average") # Clustering coefficient
```

<br>

**11) In equation (3), page 7, the authors recall Barrat et al’s clustering coefficient for weighted networks. I also introduced this index in my lectures, but equation (3) and my definition do not define the same number. What is the difference? Which one corresponds to Barrat et al’s original definition? How did you check it? If the correct version is the one given in this paper, I probably had a good reason to change it. What was this reason? And, if the correct version is the one I gave in the lectures, did the authors of this paper have a good reason to change it (for instance, to compare it with their proposed coefficient)?**


<br>

**12) In equations (4)–(5), page 8, the authors recall Zhang—Horvath’s weighted clustering coefficient for gene co-expression networks. Does it fit with the general philosophy of Kalma–Highman’s indices or Opshal–Panzarasa’s indices as explained in my lectures?**


<br>

**13) (Extra bonus) The authors mention that they used R to implement their coefficient and perform their experiments, but they do not provide their implementation. Implement in R their generalized clustering coefficient as well as your proposals in points (7) and (8), and repeat their experiments with your functions, or perform new experiments (for instance, using other networks, using your bounds in point (6), checking point (9), ...).**


<br>
