---
title: "Social and economic networks"
subtitle: "Handout 7"
author: "Juan José Martín y Christian Strasser"
output:
  html_document:
    number_sections: false
    highlight: tango
    toc: yes
    df_print: paged
editor_options: 
  chunk_output_type: console
---



<style type="text/css">

  body {
    background-color: #f6f7fd; 
  }
  
  a:link {
    color: #0174DF;
  }
  
  code.r {
    font-size: 14px;
  } 
  
  div pre {
    background-color:#E0ECF8;
  }
  pre {
    font-size: 14px 
  }
 
  p {
    text-align: justify;
  }
 
    
  h1, h2, h3, h4, h5, h6 {
    color: #737aaa;
  }

  th {  
    background-color:#737aaa;
    color: #FAFAFA;
    padding:5px;
  }
  
  td {
    font-size: 11.5pt;
  } 
  
  tr:nth-child(even){
    background: white;
  }
  
  tr:nth-child(odd){ 
    background-color: #EFF8FB;
  }
</style>

***  


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=8, fig.align="center", echo=TRUE, warning=FALSE, message=FALSE,autodep = TRUE,cache=TRUE)
library(knitr)
library(igraph)
```


**1)** The file **euroroad.txt** contains the links of the international E-road network (https://en.wikipedia.org/wiki/International_E-road_network) between the European cities given in the file **euroroad_city_names.txt**. Each node in this network represents a city and an edge between two nodes denotes that they are connected by an E-road without any other city in-between. 

```{r}
df = read.csv("datasets/euroroad.txt", sep = " ")
df_n = read.csv("datasets/euroroad_city_names.txt", sep = "")
g = graph_from_data_frame(df, directed = TRUE)
```

<br>

*a)* Provide a statistical summary of it. What are its most central cities?

Datos estadísticos del grafo:

* Order: `r igraph::gorder(g)`. 

* Tamaño: `r igraph::gsize(g)`

* Densidad: `r round(igraph::edge_density(g), 4)`

* Componentes: El número de componentes conexos es `r igraph::components(g)$no` y sus tamaños son `r igraph::components(g)$csize`

* Diámetro: `r igraph::diameter(g)`

* Distancia media: `r mean_distance(g)`

* Transitividad: `r round(igraph::transitivity(g), 4)`

<br>

Obtenemos las ciudades más centrales mediante la centralidad Betweeness de la siguiente forma:

```{r}
r = round(gorder(g) * 0.01)
c_names = names(round(igraph::betweenness(g),4)[1:r])
most_central = c()
for(i in c_names){
  most_central = c(most_central, as.character(df_n[i,1]))
}
most_central
```

<br>

De esta manera se muestra las ciudades por donde se ha de pasar más para llegar a las demás ciudades.

<br>

*b)* How do you think a road network is formed?


<br>

*c)* Is this road network compatible with the E-R model, the configuration model, or the basic Barabasi-Albert's model? Justify your answer, and comment whether you expected it.

Obtener orden, tamaño y densidad de $g$, necesario para las simulaciones posteriores, donde generaremos 1000 veces un grafo aleatorio mediante los modelos de E-R, de configuración y de Barabasi-Albert.

```{r}
order_g = gorder(g)
size_g = gsize(g)
den_g = edge_density(g, loops = FALSE)
```

<br>

E-R Model:

```{r}
ER = function(x,y) {
  G = sample_gnp(x, y, directed = TRUE, loops = FALSE)
  c(transitivity(G, type="average"), transitivity(G), components(G)$no, diameter(G))
}

X1 = replicate(1000, ER(order_g, den_g))
Results1 = rbind(c(transitivity(g, type="average"), transitivity(g), components(g)$no, diameter(g)),
               round(apply(X1, FUN = mean, MARGIN = 1), 4),
               round(apply(X1, FUN = sd, MARGIN = 1), 4),
               round(apply(X1, FUN = max, MARGIN = 1), 4))
dimnames(Results1) = list(c("Real", "Simulated", "St. Dev.", "Max."), c("Av. Clust.", "Trans.", "No. Conn. Comp.", "Diameter"))
Results1
```


<br>

Después de generar 1000 veces un grafo aleatorio mediante el modelo E-R, a partir del tamaño y la densidad del grafo de nuestra red, visualizamos los diferentes valores que nos interesan entre el grafo "real" y el grafo aleatorio "simulado". Así podemos ver que con una simulación de 1000 iteraciones, no se acerca mucho al grafo original. Dado que el modelo genera todos los enlaces con la misma probabilidad, la probabilidad de generar el grafo "real" únicamente a partir del tamaño y la densidad no es muy elevado. Por lo tanto, diremos que no cumple con un modelo E-R.

<br>

Configuration Model:

```{r}
CM = function(x,y){
  G = sample_degseq(out.deg = x, in.deg = y, method = "simple")
  G = simplify(G)
  c(transitivity(G, type="average"), transitivity(G), components(G)$no, diameter(G))
}

X2 = replicate(1000, CM(degree(g, mode = "out"), degree(g, mode = "in")))
Results2 = rbind(c(transitivity(g, type="average"), transitivity(g), components(g)$no, diameter(g)),
               round(apply(X2, FUN = mean, MARGIN = 1), 4),
               round(apply(X2, FUN = sd, MARGIN = 1), 4),
               round(apply(X2, FUN = max, MARGIN = 1), 4))
dimnames(Results2) = list(c("Theoretical", "Simulated", "St. Dev.", "Max."), c("Av. Clust.", "Trans.", "No. Conn. Comp.", "Diameter"))
Results2
```

<br>

Ahora vemos el resultado de la simulación del modelo de configuración que, a partir de la secuencia de grados de entrada y salida de los nodos de nuestro grafo, nos genera grafos aleatorios que se acerca en sus valores al grafo "real". Igual que el modelo anterior, no se acerca del todo a los valores del grafo real, por lo tanto, diremos que no cumple con el modelo de configuración.

<br>

Barabasi-Albert Model:

```{r}
BAM = function(x, y) {
  G = sample_pa(n = x, m = y, directed = TRUE, out.pref = TRUE)
  c(transitivity(G, type = "average"), transitivity(G), components(G)$no, diameter(G))
}

X3 = replicate(1000, BAM(order_g, 3))
Results3 = rbind(c(transitivity(g, type = "average"), transitivity(g), components(g)$no, diameter(g)), 
                 round(apply(X3, FUN = mean, MARGIN = 1), 4),
                 round(apply(X3, FUN = sd, MARGIN = 1), 4),
                 round(apply(X3, FUN = max, MARGIN = 1), 4))
dimnames(Results3) = list(c("Real", "Simulated", "St. Dev.", "Max."), c("Av. Clust.", "Trans.", "No. Conn. Comp.", "Diameter"))
Results3
```

<br>

También diremos que nuestra red no cumple con el modelo de Babarasi-Albert porque la probabilidad de generar el grafo real a partir del modelo de Barabasi-Albert es muy bajo.

<br>

*d)* If we add an initial attractiveness $a$ to all nodes, so that the probability of an old node $v$ to be linked by a new node is proportional to $a+\deg(v)$, the absolute value of the exponent in the power law describing its degree distribution turns out to be $3+a/c$ (with $c$ the mean number of links of the new nodes). This exponent can be estimated with **igraph**'s function **fit_power_law**. 

Now, is the E-road network compatible with Barabasi-Albert's model with initial attractiveness? Justify your answer, and comment whether you expected it.

```{r}
alpha = fit_power_law(degree(g), xmin = 3)$alpha
alpha
```


```{r}
BAM_attr = function(x, y) {
  G = sample_pa(n = x, m = y, directed = TRUE, out.pref = TRUE, zero.appeal = alpha)
  c(transitivity(G, type = "average"), transitivity(G), components(G)$no, diameter(G))
}

X4 = replicate(1000, BAM_attr(order_g, 3))
Results4 = rbind(c(transitivity(g, type = "average"), transitivity(g), components(g)$no, diameter(g)), 
                 round(apply(X4, FUN = mean, MARGIN = 1), 4),
                 round(apply(X4, FUN = sd, MARGIN = 1), 4),
                 round(apply(X4, FUN = max, MARGIN = 1), 4))
dimnames(Results4) = list(c("Real", "Simulated", "St. Dev.", "Max."), c("Av. Clust.", "Trans.", "No. Conn. Comp.", "Diameter"))
Results4
```




<br>

**2)** Consider the WikiVote directed network introduced in the Handout 4.

```{r}
df = read.csv("datasets/WikiVote.txt", sep = "\t")
g = graph_from_data_frame(df, directed = TRUE)
```


*a)* Is it compatible with the directed E-R model, the directed configuration model, or the  directed Barabasi-Albert's model (either basic or with initial attractiveness)? Justify your answer, and comment whether you expected it.


```{r}
order_g = gorder(g)
size_g = gsize(g)
den_g = edge_density(g, loops = FALSE)
```

<br>

Probamos, mediante simulaciones de los 3 modelos, si la red es compatible con alguno de ellos, tal como hemos hecho en el ejercicio 1.

E-R Model:

```{r}
ER = function(x,y) {
  G = sample_gnp(x, y, directed = TRUE, loops = FALSE)
  c(transitivity(G, type="average"), transitivity(G), components(G)$no, diameter(G))
}

X1 = replicate(10, ER(order_g, den_g))
Results1 = rbind(c(transitivity(g, type="average"), transitivity(g), components(g)$no, diameter(g)),
               round(apply(X1, FUN = mean, MARGIN = 1), 4),
               round(apply(X1, FUN = sd, MARGIN = 1), 4),
               round(apply(X1, FUN = max, MARGIN = 1), 4))
dimnames(Results1) = list(c("Real", "Simulated", "St. Dev.", "Max."), c("Av. Clust.", "Trans.", "No. Conn. Comp.", "Diameter"))
Results1
```


<br>


Configuration Model:

```{r}
CM = function(x,y){
  G = sample_degseq(out.deg = x, in.deg = y, method = "simple")
  G = simplify(G)
  c(transitivity(G,type = "average"), transitivity(G), components(G)$no, diameter(G))
}

X2 = replicate(20, CM(degree(g, mode = "out"), degree(g, mode = "in")))
Results2 = rbind(c(transitivity(g, type="average"), transitivity(g), components(g)$no, diameter(g)),
               round(apply(X2, FUN = mean, MARGIN = 1), 4),
               round(apply(X2, FUN = sd, MARGIN = 1), 4),
               round(apply(X2, FUN = max, MARGIN = 1), 4))
dimnames(Results2) = list(c("Theoretical", "Simulated", "St. Dev.", "Max."), c("Av. Clust.", "Trans.", "No. Conn. Comp.", "Diameter"))
Results2
```


<br>


Barabasi-Albert Model:

```{r}
BAM = function(x, y) {
  G = sample_pa(n = x, m = y, directed = TRUE, out.pref = TRUE)
  c(transitivity(G, type = "average"), transitivity(G), components(G)$no, diameter(G))
}

X3 = replicate(100, BAM(order_g, 3))
Results3 = rbind(c(transitivity(g, type = "average"), transitivity(g), components(g)$no, diameter(g)), 
                 round(apply(X3, FUN = mean, MARGIN = 1), 4),
                 round(apply(X3, FUN = sd, MARGIN = 1), 4),
                 round(apply(X3, FUN = max, MARGIN = 1), 4))
dimnames(Results3) = list(c("Real", "Simulated", "St. Dev.", "Max."), c("Av. Clust.", "Trans.", "No. Conn. Comp.", "Diameter"))
Results3
```

<br>


*b)* It should be clear that this network does not "grow" and that the number of votes received by each individual is somehow related to some notion of "fitness". **igraph** provides some functions to generate random networks with these features (based on the model explained in https://arxiv.org/pdf/cond-mat/0106565.pdf). Is this WikiVote network compatible with the 
"Scale-free random graphs with vertex fitness scores" provided by the function **sample_fitness_pl**? Justify and comment your answer.


```{r}

```





