---
title: "Social and economic networks"
subtitle: "Handout 4.2"
author: "Juan José Martín y Christian Strasser"
output:
  html_document:
    number_sections: false
    highlight: tango
    toc: yes
    df_print: paged
editor_options: 
  chunk_output_type: console
---



<style type="text/css">

  body {
    background-color: #f6f7fd; 
  }
  
  a:link {
    color: #0174DF;
  }
  
  code.r {
    font-size: 14px;
  } 
  
  div pre {
    background-color:#E0ECF8;
  }
  pre {
    font-size: 14px 
  }
 
  p {
    text-align: justify;
  }
 
    
  h1, h2, h3, h4, h5, h6 {
    color: #737aaa;
  }

  th {  
    background-color:#737aaa;
    color: #FAFAFA;
    padding:5px;
  }
  
  td {
    font-size: 11.5pt;
  } 
  
  tr:nth-child(even){
    background: white;
  }
  
  tr:nth-child(odd){ 
    background-color: #EFF8FB;
  }
</style>

***  

```{r setup, include=FALSE}
library(igraph)
```


## Ejercicio 2

**The file WikiVote.txt is a data table containing the votations on promotion to administratorship of Wikipedia (till January 2008). Each arc $(x, y)$ in it means that user $x$ voted on $y$ becoming Wikipedia administrator. More details at** http://snap.stanford.edu/data/wiki-Vote.html

**Download it on your computer and upload it to R as a dataframe. Define a directed graph $DW$ with this list of arcs, and the underlying unweighted undirected graph $UW$.**

```{r, results='hide'}
df = read.csv("datasets/WikiVote.txt", sep="\t")
dw = graph_from_data_frame(df, directed = TRUE)
uw = graph_from_data_frame(df, directed = FALSE)
```

<br>

**(a) Does $DW$ contain self-loops? How many? What do self-loops mean, in the context of this graph? If any, remove them (as well as from $UW$).**

```{r}
sum(which_loop(dw))
```

<br>

$DW$ no contiene ningún self-loop. No obstante, en el contesto de este grafo, un self-loop sería una arista que tiene como nodo de salia y como nodo de entrada al mismo.

```{r}
is_simple(uw)
```

<br>

Vemos que $uw$ no es simple, por lo que contiene aristas múltiple o aristas loop, que eliminamos a continuación:

```{r}
uw = simplify(uw, remove.multiple = TRUE, remove.loops = TRUE, edge.attr.comb = igraph_opt("edge.attr.comb"))
is_simple(uw)
```

<br>

Y ya podemos observar como es simple.

<br>

**(b) What are the order and size of $DW$? What is the size of $UW$? What does the difference in sizes correspond to?**

Order y tamaño de $dw$:
```{r}
gorder(dw)
gsize(dw)
```

<br>

Tamaño de $uw$:
```{r}
gsize(uw)
```

<br>

Esta diferencia se corresponde a la cantidad de aristas eliminadas de $uw$ que eran multiple o loop.

<br>

**(c) Does $DW$ have a giant weakly connected component? And a giant strongly connected component?**

```{r}
gorder(induced_subgraph(dw, v=groups(components(dw, mode = "weak"))[[1]]))
```

<br>

Sí tiene un component conexo gigante débilmente conectado de 7066 nodos.

```{r}
gorder(induced_subgraph(dw, v=groups(components(dw, mode = "strong"))[[1]]))
```

<br>

No tiene un component conexo gigante fuertemente conectado ya que el orden es 1.

<br>

**(d) Plot its out-degree distribution on a log-log scale (omitting the nodes with out-degree 0).**

```{r, fig.align='center'}
dg_dist_out = degree.distribution(dw, mode="out")
dg_dist_out = dg_dist_out[dg_dist_out != 0]

plot(dg_dist_out, log='xy')
```

<br>

**(e) Do the same with its in-degree distribution.**

```{r, fig.align='center'}
dg_dist_in = degree.distribution(dw, mode="in")
dg_dist_in = dg_dist_in[dg_dist_in != 0]

plot(dg_dist_in, log='xy')
```

<br>

**(f) For both the in-degrees and the out-degrees, compute their range, mode, mean, median, and standard deviation. What do they tell you?**


Rango OUT:
```{r}
range(dg_dist_out)
```

<br>

Moda OUT:
```{r}
names(sort(table(degree(dw, mode = "out")), decreasing = TRUE))[1]
```

<br>

Media OUT:
```{r}
mean(dg_dist_out)
```

<br>

Mediana OUT:
```{r}
median(dg_dist_out)
```

<br>

Desviación estándar OUT:
```{r}
sd(dg_dist_out)
```

<br>
<br>


Rango IN:
```{r}
range(dg_dist_in)
```

<br>

Moda IN:
```{r}
names(sort(table(degree(dw, mode = "in")), decreasing = TRUE))[1]
```

<br>

Media IN:
```{r}
mean(dg_dist_in)
```

<br>

Mediana IN:
```{r}
median(dg_dist_in)
```

<br>

Desviación estándar IN:
```{r}
sd(dg_dist_in)
```

<br>

**(g) What kind (or kinds) of transitivity do you consider worth studying in this network? Choose at least one, compute it, and comment the results.**

Consideramos oportuno estudiar la transitividad de doble-entrada para medir cuántas veces, por ejemplo, si $X$ e $Y$ muestran estar conectados a $Z$, implica que también deben estar conectados entre sí.

Tipo de tripleta '15': A->B<->C, A<->C: 2 1

Función para transitividad de doble-entrada:

```{r}
Local_Double_In_Trans = function(G, x){
  NN = induced_subgraph(G, v = as.vector(neighbors(G, v = x, mode = "in")))
  (sum(which_mutual(NN)) / 2) / choose(gorder(NN), 2)
}
```

<br>

Transitividad de doble-entrada del grafo dirigido _DW_:

```{r}
Local_Double_In_Trans(dw, 15)
```

<br>

Como podemos ver, se trata de un valor muy bajo, por lo que podemos considerar que tendremos más nodos con grado de entrada muy grandes y cuyos nodos que entran en ellos no estarán conectados entre sí. 

<br> 