---
title: "Social and economic networks"
subtitle: "Handout 4.A"
author: "Juan José Martín y Christian Strasser"
output:
  html_document:
    number_sections: false
    highlight: tango
    toc: yes
    df_print: paged
editor_options: 
  chunk_output_type: console
---



<style type="text/css">

  body {
    background-color: #f6f7fd; 
  }
  
  a:link {
    color: #0174DF;
  }
  
  code.r {
    font-size: 14px;
  } 
  
  div pre {
    background-color:#E0ECF8;
  }
  pre {
    font-size: 14px 
  }
 
  p {
    text-align: justify;
  }
 
    
  h1, h2, h3, h4, h5, h6 {
    color: #737aaa;
  }

  th {  
    background-color:#737aaa;
    color: #FAFAFA;
    padding:5px;
  }
  
  td {
    font-size: 11.5pt;
  } 
  
  tr:nth-child(even){
    background: white;
  }
  
  tr:nth-child(odd){ 
    background-color: #EFF8FB;
  }
</style>

***  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Ejercicio 1

**The file facebook_sample_anon.txt is a data table containing the list of edges of an anonymized sample of the Facebook friendship network. Download it on your computer, upload it to R as a dataframe, and define an undirected graph with this list of edges.**

```{r}
library(igraph)
df = read.csv("datasets/facebook_sample_anon.txt", sep=" ")

g = graph_from_data_frame(df, directed = FALSE)
```

<br>

(a) **What are its order and size?**

Orden:
```{r}
gorder(g)
```

<br>

Tamaño:
```{r}
ecount(g) 
```

<br>

(b) **Is it connected?**

Descomponer en subgrafos. Es conexo si solo hay 1 subgrafo (el máximo).

```{r}
length(decompose(g))
```

<br>

(c) **Compute its density. Do you consider it sparse?**

```{r}
edge_density(g, loops = FALSE)
```

<br>

Es un grafo disperso porque la densidad del grafo es de un 1%.

<br>

(d) **Plot its degrees distribution in linear and in log-log scale. Which is more helpful to understand this distribution? Does this distribution seem to follow a power-law?**

```{r}
dg_dist = degree.distribution(g)
plot(dg_dist)
plot(dg_dist, log='xy')
```

<br>

(e) **What are the range, the mode, the mean, the median and the standard deviation of its degree distribution? What do they tell you?**

```{r}
range(dg_dist)
names(sort(table(degree(g)), decreasing = TRUE))[1]
mean(dg_dist)
median(dg_dist)
sd(dg_dist)
```

<br>

(f) **Compute its diameter, its average distance, and its 90% effective diameter (the smallest distance d such that at least a 90% of the pairs of different nodes are at distance at most d).**

Diámetro:
```{r}
diameter(g, directed = FALSE)
```

<br>

Distancia media:
```{r}
round(mean_distance(g))
```

<br>

Distancia a la que están el 90% de los nodos: 
```{r}
l = sort(distances(g),decreasing = FALSE)
l[round(length(l) * 0.9)]
```

<br>

(g) **Plot a histogram of its clustering coefficients.**

```{r}
hist(round(transitivity(g, type="barrat"), 3), main="Clustering Coefficients", xlab="Probability", ylab="Num. of Nodes")
```

<br>

(h) **Compute its average clustering coefficient and its transitivity coefficient. What do they tell you about the clustering coefficient of the nodes with large degree? Confirm it by computing the clustering coefficients of the 10 nodes with largest degree.**

Coeficiente de clustering:
```{r}
round(transitivity(g, type="average"),4)
```

<br>

Coeficiente de transitividad:
```{r}
round(transitivity(g),4)
```

<br>

El coeficiente de transitividad es menor que el coeficiente de clustering, lo que significa que la probabilidad es muy baja de que dos nodos est?n conectados entre si, cuando est?n conectados a un nodo con grado alto.

Se puede demostrar cogiendo solo el conjunto de los 10 nodos con mayor grado y obtener la media de su coeficiente de clustering y obtener un valor relativamente bajo.
```{r}
coef_clus=0
media_nodos_mayor = c()
conjunto_nodos_mayor = names(sort(degree(g), decreasing = TRUE))[1:10]
for(i in 1:10){
  media_nodos_mayor = c(media_nodos_mayor, transitivity(g, type="local", vids = conjunto_nodos_mayor[i]))
}
mean(media_nodos_mayor)
```

<br>

(i) **This graph was obtained by gathering all friends of 10 students enrolled in a course and establishing the friendship relations among the union of their friends (in the old times when Facebook APIs allowed it easily). Can you make an educated guess about which are these 10 students?**

Serían los 10 nodos que no están conectados directamente entre ellos y donde 2 otros nodos cualquieras tienen un camino entre ellos que pasa por uno de estos 10 nodos iniciales.

Los 10 nodos con grados más grandes que se encuentran en el camino de dos otros nodos cualquieras.
```{r}

```