---
output:
  html_document: default
  pdf_document: default
  word_document: default
---

<style type="text/css">
    body {
        background-color: #f6f7fd;
    }
    
    p {
        text-align: justify;
        text-indent: 1cm;
    }
    
    li p {
        text-indent: 0cm;
    }
    
    h1, h2, h3, h4, h5, h6 {
        color: #737aaa;
        text-decoration: underline;
    }
    
    th {  
        background-color:#737aaa;
        color: white;
        padding:5px;
    }
    
    tr:nth-child(even){
		    background: #dae5f4;
	  }
	  
	  tr:nth-child(odd){ 
		    background: #b8d1f3;
	  }
</style>

---
title: "Hoja de ejercicios 2"
author: "Author name"
output: html_document
---
    
***       

<br/>    

### Problema 1

Enunciado

<br/>   

```{r}
```      

<br/>    

### Problema 2

Enunciado

<br/>   

```{r}
```      

<br/>    

### Problema 3

Enunciado

<br/>   

```{r}
```      

<br/>    

### Problema 4

Enunciado

<br/>   

```{r}
```      

<br/>    

### Problema 5

Enunciado

<br/>   

```{r}
```      

<br/>    

### Problema 6

Enunciado

<br/>   

```{r}
```

